\documentclass{article}
\usepackage{booktabs}
\usepackage{geometry}
\usepackage{tabularx}
\usepackage{array}
\setlength{\extrarowheight}{2pt}
\geometry{margin=1in}

\begin{document}

\section*{Architectural Comparison of Proposed Local GGUF Models}

When executing LLM inference on pure CPU infrastructure, performance is heavily bound by memory bandwidth rather than raw FLOPS. The tables below outline the architectural trade-offs, pros, and cons of each proposed configuration quantized at the $Q4\_K\_M$ level.

\noindent
\small
\begin{tabularx}{\textwidth}{l p{2.0cm} c >{\raggedright\arraybackslash}X >{\raggedright\arraybackslash}X}
\toprule
\textbf{Model Family} & \textbf{Effective Params} & \textbf{RAM Required} & \textbf{Primary Strengths} & \textbf{Compute Footprint} \\
\midrule
\textbf{Llama 3.1 (8B)} & 8.03 Billion & $\sim$4.8 GB & General reasoning, canonical tool-use instruction following. & Moderate-high memory bandwidth saturation. \\
\textbf{Qwen 2.5 (7B)} & 7.61 Billion & $\sim$4.7 GB & Code generation, structured JSON extraction, multilingual math. & Dense attention matrix operations. \\
\textbf{Qwen 2.5 Coder (1.5B)} & 1.50 Billion & $\sim$1.0 GB & Extremely low-latency tab-autocomplete and IDE-assist; very small memory footprint for always-on services. & Minimal KV-cache and low memory bandwidth demands. \\
\textbf{Phi-3.5-Mini} & 3.82 Billion & $\sim$2.6 GB & Low-latency execution, minimal memory footprint. & Lightweight tracking paths; fast prompt ingestion. \\
\textbf{Gemma 2 (9B)} & 9.00 Billion & $\sim$5.4 GB & Strong instruction-following, balanced reasoning, and code capability. & Higher memory bandwidth pressure with a slightly larger cache footprint. \\
\textbf{Mistral 7B Instruct} & 7.24 Billion & $\sim$4.2 GB & Fast general-purpose chat, clean instruction-following, and reliable summarization. & Efficient decoder stack with modest KV-cache demands. \\
\textbf{Qwen 2.5 (14B)} & 14.78 Billion & $\sim$8.6 GB & Better reasoning depth, stronger code synthesis, and robust structured output. & Heavier attention and feed-forward passes, but still comfortable in 128GB RAM. \\
\textbf{Llama 3.1 (70B)} & 70.55 Billion & $\sim$40 GB & High-quality reasoning, long-form planning, and strong tool orchestration. & Large weight footprint; best used with conservative context sizes on CPU. \\
\bottomrule
\end{tabularx}

\vspace{0.5cm}

\subsection*{1. Meta Llama 3.1 (8B Instruct)}
\begin{itemize}
    \item \textbf{Pros:}
    \begin{itemize}
        \item Highly stable attention mechanics with an expanded native context window up to 128K tokens (though performance scales down with token depth on standard RAM configurations).
        \item Extremely reliable at maintaining character/agent persona boundaries over long local runtime sessions.
        \item Strong multi-turn dialogue management with clear system prompt adherence.
    \end{itemize}
    \item \textbf{Cons:}
    \begin{itemize}
        \item High prompt processing overhead (time-to-first-token) when digesting large context blocks over pure CPU execution matrices.
        \item Sacrifices specialized code-generation capabilities compared to hyper-dense domain-specific alternatives of equal size.
    \end{itemize}
\end{itemize}

\noindent\textbf{Download command example:}
\begin{verbatim}
from huggingface_hub import hf_hub_download

hf_hub_download(
    repo_id="bartowski/Meta-Llama-3.1-8B-Instruct-GGUF",
    filename="Meta-Llama-3.1-8B-Instruct-Q4_K_M.gguf",
    local_dir=".",
    token="Insert_Token_Here"
)
\end{verbatim}

\subsection*{2. Alibaba Qwen 2.5 (7B Instruct)}
\begin{itemize}
    \item \textbf{Pros:}
    \begin{itemize}
        \item Phenomenal performance on system-level tasks: syntax parsing, writing regular expressions, processing structural logs, and script generation.
        \item Highly optimized vocabulary encoding ($\sim$151k tokens), which allows it to express more compressed data per token slice than Llama or Phi.
        \item Natively structured to emit deterministic JSON schemas when wrapped with strict local grammar constraints (using \texttt{--json-schema} configurations in \texttt{llama.cpp}).
    \end{itemize}
    \item \textbf{Cons:}
    \begin{itemize}
        \item Large vocabulary size shifts a larger portion of the model's weight tensor allocation into the embedding layer, slightly increasing structural RAM overhead per parameter.
        \item Can occasionally over-generate conversational filler text unless explicitly constrained by strict system stop-tokens.
    \end{itemize}
\end{itemize}

\noindent\textbf{Download command example:}
\begin{verbatim}
from huggingface_hub import hf_hub_download

hf_hub_download(
    repo_id="bartowski/Qwen2.5-7B-Instruct-GGUF",
    filename="Qwen2.5-7B-Instruct-Q4_K_M.gguf",
    local_dir=".",
    token="Insert_Token_Here"
)
\end{verbatim}

\subsection*{2a. Alibaba Qwen 2.5 Coder (1.5B — Tab Autocomplete)}
\begin{itemize}
    \item \textbf{Pros:}
    \begin{itemize}
        \item Extremely low-latency and compact for IDE tab-autocomplete and inline code suggestions.
        \item Small memory footprint makes it suitable for always-on background completion services on developer machines or constrained servers.
        \item Retains strong tokenization and deterministic completions when constrained with strict stop sequences.
    \end{itemize}
    \item \textbf{Cons:}
    \begin{itemize}
        \item Lower overall reasoning and long-form synthesis capability compared to 7B+ models.
        \item May require careful temperature and sampling tuning to avoid repetitious completions in interactive sessions.
    \end{itemize}
\end{itemize}

\noindent\textbf{Download command example:}
\begin{verbatim}
from huggingface_hub import hf_hub_download

hf_hub_download(
    repo_id="bartowski/Qwen2.5-coder-1.5B-GGUF",
    filename="Qwen2.5-Coder-1.5B.Q4_K_M.gguf",
    local_dir=".",
    token="Insert_Token_Here"
)
\end{verbatim}


\subsection*{3. Microsoft Phi-3.5-Mini (Instruct)}
\begin{itemize}
    \item \textbf{Pros:}
    \begin{itemize}
        \item Blistering fast token generation rates (tokens-per-second) on standard server hardware without GPU acceleration.
        \item Highly optimized for processing simple local data streams, local indexing tasks, or voice-to-text semantic cleaning pipelines.
        \item Massive parameter efficiency; matches older 7B and 13B parameter architectures on logical benchmarks due to highly curated training datasets.
    \end{itemize}
    \item \textbf{Cons:}
    \begin{itemize}
        \item Deep multi-step reasoning loops or complex code debugging sequences degrade significantly faster due to the lower raw parameter ceiling.
        \item Prone to lower semantic density limits when analyzing long, multi-layered log files or wide text corpora simultaneously.
    \end{itemize}

\end{itemize}

\noindent\textbf{Download command example:}
\begin{verbatim}
from huggingface_hub import hf_hub_download

hf_hub_download(
    repo_id="bartowski/Phi-3.5-mini-instruct-GGUF",
    filename="Phi-3.5-mini-instruct-Q4_K_M.gguf",
    local_dir=".",
    token="Insert_Token_Here"
)
\end{verbatim}

\subsection*{4. Google Gemma 2 (9B Instruct)}
\begin{itemize}
    \item \textbf{Pros:}
    \begin{itemize}
        \item Strong instruction-following and balanced general reasoning at a size that remains friendly to pure CPU inference.
        \item Good code completion and editing behavior for local development workflows.
        \item Efficient enough to leave ample headroom for larger context windows and concurrent OS services on a 128GB host.
    \end{itemize}
    \item \textbf{Cons:}
    \begin{itemize}
        \item More memory-hungry than 7B-class models, so throughput drops sooner if context windows are pushed aggressively.
        \item Can be less specialized for structured extraction than Qwen-family models.
    \end{itemize}
\end{itemize}

\noindent\textbf{Download command example:}
\begin{verbatim}
from huggingface_hub import hf_hub_download

hf_hub_download(
    repo_id="bartowski/gemma-2-9b-it-GGUF",
    filename="gemma-2-9b-it-Q4_K_M.gguf",
    local_dir=".",
    token="Insert_Token_Here"
)
\end{verbatim}

\subsection*{5. Mistral 7B Instruct}
\begin{itemize}
    \item \textbf{Pros:}
    \begin{itemize}
        \item Very strong baseline for a compact local model: good instruction following, concise answers, and fast inference.
        \item Excellent fit for always-on CPU services where latency and simplicity matter more than maximum reasoning depth.
        \item Leaves significant RAM headroom for long prompts, retrieval buffers, and parallel background processes.
    \end{itemize}
    \item \textbf{Cons:}
    \begin{itemize}
        \item Less capable than larger 9B-14B models on harder reasoning and coding tasks.
        \item Can be somewhat conservative or terse unless prompted carefully.
    \end{itemize}
\end{itemize}

\noindent\textbf{Download command example:}
\begin{verbatim}
from huggingface_hub import hf_hub_download

hf_hub_download(
    repo_id="bartowski/Mistral-7B-Instruct-v0.3-GGUF",
    filename="Mistral-7B-Instruct-v0.3-Q4_K_M.gguf",
    local_dir=".",
    token="Insert_Token_Here"
)
\end{verbatim}

\subsection*{6. Qwen 2.5 (14B Instruct)}
\begin{itemize}
    \item \textbf{Pros:}
    \begin{itemize}
        \item Stronger reasoning and code synthesis than the 7B variant while preserving excellent structured-output behavior.
        \item Very practical on a 128GB dual-CPU server, especially if the workload mixes coding, logs, and tool use.
        \item Good balance of quality and resource usage for serious local assistant workflows.
    \end{itemize}
    \item \textbf{Cons:}
    \begin{itemize}
        \item Heavier than the 7B-class models and slower at the same quantization level.
        \item More sensitive to context length growth and KV-cache size when running multiple long sessions.
    \end{itemize}
\end{itemize}

\noindent\textbf{Download command example:}
\begin{verbatim}
from huggingface_hub import hf_hub_download

hf_hub_download(
    repo_id="bartowski/Qwen2.5-14B-Instruct-GGUF",
    filename="Qwen2.5-14B-Instruct-Q4_K_M.gguf",
    local_dir=".",
    token="Insert_Token_Here"
)
\end{verbatim}

\subsection*{7. Meta Llama 3.1 (70B Instruct)}
\begin{itemize}
    \item \textbf{Pros:}
    \begin{itemize}
        \item Best suited for the most demanding local reasoning, planning, and long-form synthesis tasks.
        \item A strong use case for a 128GB server if you want one flagship model rather than several smaller specialists.
        \item Excellent quality at moderate quantization levels, with enough RAM available to support the model plus runtime overhead.
    \end{itemize}
    \item \textbf{Cons:}
    \begin{itemize}
        \item Much slower than smaller models on pure CPU hardware, especially for first-token latency.
        \item Requires careful context budgeting and may reduce concurrency if the machine is shared with other services.
    \end{itemize}
\end{itemize}

\noindent\textbf{Download command example:}
\begin{verbatim}
from huggingface_hub import hf_hub_download

hf_hub_download(
    repo_id="bartowski/Meta-Llama-3.1-70B-Instruct-GGUF",
    filename="Meta-Llama-3.1-70B-Instruct-Q4_K_M.gguf",
    local_dir=".",
    token="Insert_Token_Here"
)
\end{verbatim}

\end{document}